NMRlipids Databank makes data-driven analysis of biomembrane properties accessible for all

### 2023-03-02 main.tex
Update text and pdf

The focus of biomolecular simulations is moving from studies of individual molecules to larger complexes and even whole cells and organelles~\cite{johnson15,thornburg22,gupta22}. Simultaneously, machine learning based models predicting behaviour of biomolecules and automatic approaches to parametrize models are emerging~\cite{jumper21,antila22b}. The NMRlipids databank will support the development in both of these directions. The automatic quality evaluation and ranking of simulations in the NMRlipids databank guides the optimization of new force field parameters and the selection of best parameters for large biomolecular complexes where simulation quality becomes increasingly important due the accumulation of small errors in large simulations. Furthermore, the NMRlipids databank serves as a training set for diverse machine learning applications. For example, a machine learning model predicting electron density profiles from form factors would be highly useful in interpretation of scattering experiments. In addition, the data contained in the NMRlipids databank will enable training of more elaborated models to predict arbitrary membrane properties.

These tools offer open programmatic access to unprecedented amount of simulation trajectories for all, which enables possibilities for wide range of novel analyses. This is particularly useful for scientist without access to resources to perform large scale MD simulations and for analyses of rare phenomena that require exceptional amounts of simulation data. On the other hand, universal naming convention enables systematic comparisons between simulations having different model parameters and thereby different naming conventions. This will guide the development of more accurate simulation parameters and usage of atomistic models to train coarse grained simulations.

%The NMRlipids databank will be of great benefit to scientists involved in MD simulations. They will be able to rapidly evaluate the quality of MD simulations in order to filter out potentially misleading results or facilitate force field parameter development~\cite{antila22b}. The selection of best models for a specific application is demonstrated for PC/PE lipid mixtures in Fig.~\ref{fig:quality}, and previously for PC/PS lipid mixtures~\cite{antila22b}, and lipid headgroups~\cite{bacle21}. 
The programmatic access in the NMRlipids databank enables automatic analyses of structural and dynamic properties over larger sets of simulation data in terms of quantity (\textit{e.g.}, simulation length and number of conformations) and content (\textit{e.g.}, lipid compositions and ion concentrations) than is currently possible by a single research group. The added value of such analyses is demonstrated here by analysing the anisotropic diffusion of water in membrane systems (Fig.~\ref{fig:permeability}) and cholesterol flip--flop events (Fig.~\ref{fig:flip-flops}). By enabling these analyses for everyone, the NMRlipids databank can lead to applications in unprecedented directions in fields where MD simulations are less commonly utilized, such as pharmacokinetic modeling and in MRI imaging where anisotropic diffusion of water play important roles~\cite{nitsche19,topgaard20}. 
Different types of applications enabled by the NMRlipids databank across a wide range of fields are exemplified in Table~\ref{tab:applications}. The ever growing amount of data is expected to further increase the scope of potential applications of the NMRlipids databank in biology, biotechnology, material science, biomolecular imaging and beyond.

>>>>

The focus of biomolecular simulations is moving from studies of individual molecules to larger complexes and even whole cells and organelles~\cite{johnson15,thornburg22,gupta22}. Simultaneously, machine learning based models predicting behaviour of biomolecules and automatic approaches to parametrize models are emerging~\cite{jumper21,antila22b}. The new NMRlipids databank will support the development in both of these directions. The automatic quality evaluation and ranking of simulations against experimental data in the NMRlipids databank can be used to select the best parameters for simulations of large biomolecular complexes and to guide optimizations of new force field parameters for both atomistic and coarse grained simulations. These issues are becoming increasingly important due the accumulation of small errors when complexity and size of simulated systems are increasing.

The open programmatic access to unprecedented amount of MD simulation via NMRlipids databank-API enables wide range of novel data-driven analyses and machine learning applications. For example, large amount of accessible simulation data in terms of quantity (\textit{e.g.}, simulation length and number of conformations) and content (\textit{e.g.}, lipid compositions and ion concentrations) enables analyses of rare phenomena beyond current possibilities for a single research group as demonstrated for cholesterol flip--flop events (Fig.~\ref{fig:flip-flops}) and water permeation through membranes (Fig.~\ref{fig:permeability}). This will also pave the way of applications of MD simulations in new fields as demonstrated here by analysing the anisotropic diffusion of water in membrane systems (Fig.~\ref{fig:permeability}) that is essential parameter in pharmacokinetic modeling and MRI imaging~\cite{nitsche19,topgaard20}. Furthermore, the NMRlipids databank-API paves the way for diverse machine learning applications to 
%For example, a machine learning model predicting electron density profiles from form factors would be highly useful in interpretation of scattering experiments.
%In addition, the data contained in the NMRlipids databank will enable training of more elaborated models to
predict membrane properties using NMRlipids databank as a training set. Such applications are analogous to AlphaFold \cite{jumper21} and other tools that predict proteins structures from their sequence using artificial intelligence. These possibilities are particularly valuable for scientist who do not typically have access to large scale MD simulation data.
%and for analyses of rare phenomena that require exceptional amounts of simulation data.
%On the other hand, universal naming convention enables systematic comparisons between simulations having different model parameters and thereby different naming conventions.
%
%The NMRlipids databank will be of great benefit to scientists involved in MD simulations. They will be able to rapidly evaluate the quality of MD simulations in order to filter out potentially misleading results or facilitate force field parameter development~\cite{antila22b}. The selection of best models for a specific application is demonstrated for PC/PE lipid mixtures in Fig.~\ref{fig:quality}, and previously for PC/PS lipid mixtures~\cite{antila22b}, and lipid headgroups~\cite{bacle21}. 
%
Different types of expected applications of the NMRlipids databank in wide range of fields are listed in Table~\ref{tab:applications}, yet the scope of such applications
%in biology, biotechnology, material science, biomolecular imaging and beyond
is expected to further widen with increasing amount of data in the NMRlipids databank.
